---
title: "Run CPH models"
author: "AMCalejandro"
date: "1/09/2022"
output: html_document
---


# Load libraries and source util to split genetic data into chunks
```{r}
.libPaths("/data/kronos/kronos/acarrasco/R_libs")
library(dplyr)
library(data.table)
library(tidyverse)
library(doParallel)
library(survival)
library(foreach)

source(here::here("utils/cph_run_functions.R"))
```


# Load data

Load clinical data we need to fit the model
Load PCs we need to correct for pop stratification
This process is dope independently for each cohort we had access to

```{r}
PCs <- read.table("PCA.eigenvec", sep = "", header= FALSE)
clinical <- readRDS("~DATA/MYQC_COHORT.RDS")
```


# Run chunk file creater

```{r}
# Calling the function to split the big genetic file and read it
ChunkFileCreator(Ngwas, Nchunk)
```


# Do minimal data QC and join the data

```{r}
#Select just the first 5 principal components
PCs <- PCs %>%
  select(V2:V7) %>%
  rename(IID = V2,
         PC1 = V3,
         PC2 = V4,
         PC3 = V5,
         PC4 = V6,
         PC5 = V7)

clinical <- clinical %>%
    filter(left_censored != 1 ) %>%
    mutate(ldopaDose.std = scale(timeEvent_LdopaDose),
           motorBL.std = scale(timeEvent_UPDRSIIItotal_imputed))
```


# Setup parallel backend to use many processors

```{r}
cores=detectCores()
cl <- makeCluster(25) # 25 to not overload the cluster
clusterEvalQ(cl, .libPaths("/data/kronos/kronos/acarrasco/R_libs"))
registerDoParallel(cl)
```


# Run the parallel job
```{r}
chunks <- as.vector(list.files("GeneticChunksFiles/", pattern="txt", full.names = TRUE, recursive = TRUE))
total_chunks = as.integer(length(chunks))
finalMatrix <- foreach(i=1:total_chunks, .combine=rbind) %dopar% {
         chunkMatrix = GWAS_survival(clinical = clinical, PCs = PCs, i = i) #calling a function
         chunkMatrix #Equivalent to finalMatrix = cbind(finalMatrix, chunkMatrix)
  }
```


# Syop the cluster
```{r}
stopCluster(cl)
```


# Save results

```{r}
fwrite(finalMatrix, file = "COHORT_CPHGWAS_RES.txt", col.names = TRUE, sep = "\t", quote = F)
```

